Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-08-01 Multiresolution_framework/ltx/response_to_reviews.tex

Parham Aram.

\section{Handling editor}
\begin{enumerate}
\item Whilst the paper does fit the scope of NI, I agree with R1 and R3, that explicit advantages of going to a multiresolution frame - in addition to the conceptual ones - compared to the prior work based on Gaussian basis functions needs clarification and elucidation.

\emph{Response here.}

\item  In addition, at least a ``proof of principle" inversion from empirical data would enhance the potential impact of the paper.

@@

			\item This model cries out for validation in real neuronal systems, but alas, at this point we only see convergence to synthetic data.

			\emph{The authors appreciate that the only way to truly validate the proposed framework is with real data. However, before validating the method on real data we believe that it is essential to first validate the estimation method using accepted models and synthetic data, where the true and estimated parameters can be compared. In an effort to make the method employed in the paper more explicit, we have added ``To illustrate the estimation framework, data is generated using the neural field equations incorporating modeled sensors enabling a comparison between the estimated and true parameters'' to the introduction. We plan to validate the framework on real data in future work. We have added that validating the framework on real data is required in the Conclusion.}

			\item This audience needs to have all symbols, including commonly used ones such as `equality over distribution', double font domains like R and Z, and jargon terms such as `lead field'.

			\emph{\parham{The double font domains are added to Table.1.}}

### 2012-08-01 Multiresolution_framework/ltx/manuscript.tex
4.1.3

The utility of these models for answering questions in clinical neuroscience and neurology lies in the ability to infer physiological parameters from electrophysiological data. The ability to create patient-specific neural models will contribute to our knowledge base of diseases such as epilepsy and will enable the development of new treatment strategies. Large-scale neuroinformatic efforts, such as the Virtual Brain project (www.thevirtualbrain.org \citep{Jirsa2010}), are on the way to enable patient-specific full brain modeling. This is particularly relevant to the advent of new devices that use therapeutic electrical stimulation. Stimulation strategies for devices currently operate in an open loop, where stimulation parameters are chosen using a process of trial and error. Therefore, there exists enormous potential to improve the performance of these devices using systems theory, which in turn requires a suitable dynamic model. Neural mass and neural field models are ideal candidates for estimation algorithms due to their strong links with physiology and their parsimony. \parham{The main difference between these two classes is that neural field models describe the spatiotemporal evolution of a particular variable such as firing rate, membrane voltage or synaptic activity defined as the statistical expectation over a neural population while neural mass models describe such a quantity over time only, assuming that neurons are localized at approximately the same spatial location.} It is expected that parameters of the neural fields models will be patient-specific and will, therefore, need to be inferred from data.

@@

Typically, the experimentalist only has access to the net field generated by the multi-resolution architecture of the cortex. Following this, it is natural to ask how do we best describe the multi-resolution structure of the neural system. The solution to this problem was proposed by \citet{Breakspear2005} in using a wavelet decomposition. This paper is directed at developing a framework for patient-specific multi-resolution neural field models, forming a fundamental link between data and theory. 
\section{IDE Neural Field Model}
In a small network with no time delays the single layer homogeneous neural field model, in the most general form, is described by the equation
\begin{equation}
	\label{FullDoubleIntModel} \underbrace{v\left(\mathbf{r},t\right)}_{\text{field}} =
	\int_{-\infty}^t 
	\underbrace{h\left(t - t'\right)}_{\text{synapse}} \int_\Omega
	\underbrace{w\left(\mathbf{r},\mathbf{r}'\right)}_{\text{connectivity}} 
	\underbrace{f\left( v\left( \mathbf{r}',t'\right)\right)}_{\text{firing}}
	\, \mathrm{d}\mathbf{r}'\mathrm{d}t',
\end{equation}
where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ within the spatial domain $\Omega$ is denoted by $v\left(\mathbf r,t\right)$. The function $h(t)$ describes the post-synaptic response kernel, $w\left(\mathbf{r},\mathbf{r}'\right)$ describes the connectivity kernel and $f(v(\mathbf{r},t))$ describes the relationship between the mean membrane potential and the mean firing rate. There are various forms of $h(t)$ that have been proposed in the literature, with different levels of complexity, \parham{leading to first and second order differential equations}. These are plotted in Fig.~\ref{fig:Modelcomponents}, \parham{showing finite rise time (solid line) versus instantaneous response (dotted line). The second order synaptic kernel reproduces  experimental excitatory and  inhibitory post-synaptic potentials (EPSP and IPSP) features \cite{VanRotterdam1982}}. Similarly, both the connectivity kernel and the activity function (firing rate) can take various forms. These are also shown in Fig.~\ref{fig:Modelcomponents}. The form of $w(\mathbf{r},\mathbf{r}')$ typically describes a Laplacian (exponential) or Gaussian decay in connectivity strength from any point in the field \cite{Breakspear2010,Atay2005}. By assuming the form of $h(t)$ is the same across layers (scales) the Amari style neural field model is formed, incorporating a multilayer structure into the connectivity kernel. \parham{The connectivity kernel then can be written as a sum of weighted Gaussian or Laplacian functions where the weights can be thought of as the synaptic efficiency or coupling strength where positive and negative weights describe excitation and inhibition, respectively.} In this way, the connectivity kernel typically takes the form of a Mexican hat (Gaussian shape within layer connectivity) or Wizard hat (Laplacian shape within layer) as depicted in Fig.~\ref{fig:Modelcomponents}.

@@

>>>>

The utility of these models for answering questions in clinical neuroscience and neurology lies in the ability to infer physiological parameters from electrophysiological data. The ability to create patient-specific neural models will contribute to our knowledge base of diseases such as epilepsy and will enable the development of new treatment strategies. Large-scale neuroinformatic efforts, such as the Virtual Brain project (www.thevirtualbrain.org \citep{Jirsa2010}), are on the way to enable patient-specific full brain modeling. This is particularly relevant to the advent of new devices that use therapeutic electrical stimulation. Stimulation strategies for devices currently operate in an open loop, where stimulation parameters are chosen using a process of trial and error. Therefore, there exists enormous potential to improve the performance of these devices using systems theory, which in turn requires a suitable dynamic model. Neural mass and neural field models are ideal candidates for estimation algorithms due to their strong links with physiology and their parsimony. \parham{The main difference between these two classes is that neural field models describe the spatiotemporal evolution of a particular variable such as firing rate, membrane voltage or synaptic activity defined as the statistical expectation over a neural population while neural mass models describe such a quantity over time only, assuming that neurons are localized at approximately the same spatial location.} It is expected that parameters of \parham{(these models instead neural fields models ??)} the neural fields models will be patient-specific and will, therefore, need to be inferred from data.

@@

Typically, the experimentalist only has access to the net field generated by the multi-resolution architecture of the cortex. Following this, it is natural to ask how do we best describe the multi-resolution structure of the neural system. The solution to this problem was proposed by \citet{Breakspear2005} in using a wavelet decomposition. This paper is directed at developing a framework for patient-specific multi-resolution neural field models, forming a fundamental link between data and theory. 
\section{IDE Neural Field Model}
In a small network with no time delays the single layer homogeneous neural field model, in the most general form, is described by the equation
\begin{equation}
	\label{FullDoubleIntModel} \underbrace{v\left(\mathbf{r},t\right)}_{\text{field}} =
	\int_{-\infty}^t 
	\underbrace{h\left(t - t'\right)}_{\text{synapse}} \int_\Omega
	\underbrace{w\left(\mathbf{r},\mathbf{r}'\right)}_{\text{connectivity}} 
	\underbrace{f\left( v\left( \mathbf{r}',t'\right)\right)}_{\text{firing}}
	\, \mathrm{d}\mathbf{r}'\mathrm{d}t',
\end{equation}
where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ within the spatial domain $\Omega$ is denoted by $v\left(\mathbf r,t\right)$. The function $h(t)$ describes the post-synaptic response kernel, $w\left(\mathbf{r},\mathbf{r}'\right)$ describes the connectivity kernel and $f(v(\mathbf{r},t))$ describes the relationship between the mean membrane potential and the mean firing rate. There are various forms of $h(t)$ that have been proposed in the literature, with different levels of complexity, \parham{leading to first and second order differential equations}. These are plotted in Fig.~\ref{fig:Modelcomponents}, \parham{showing finite rise time (solid line) versus instantaneous response (dotted line). The second order synaptic kernel reproduces  experimental excitatory and  inhibitory post-synaptic potentials (EPSPs and IPSPs) features \cite{VanRotterdam1982}}. Similarly, both the connectivity kernel and the activity function (firing rate) can take various forms. These are also shown in Fig.~\ref{fig:Modelcomponents}. The form of $w(\mathbf{r},\mathbf{r}')$ typically describes a Laplacian (exponential) or Gaussian decay in connectivity strength from any point in the field \cite{Breakspear2010,Atay2005}. By assuming the form of $h(t)$ is the same across layers (scales) the Amari style neural field model is formed, incorporating a multilayer structure into the connectivity kernel. \parham{The connectivity kernel then can be written as a sum of weighted Gaussian or Laplacian functions where the weights can be thought of as the synaptic efficiency or coupling strength where positive and negative weights describe excitation and inhibition, respectively.} In this way, the connectivity kernel typically takes the form of a Mexican hat (Gaussian shape within layer connectivity) or Wizard hat (Laplacian shape within layer) as depicted in Fig.~\ref{fig:Modelcomponents}.

@@

	$\mathbb{R}$&\parham{The et of all real numbers}&n.a.\\
	$\mathbb{Z}$&\parham{The set of all integers}&n.a.\\
